---
title: "TP2"
format: html
editor: visual
---

This session is dedicated to the study of a single dataset, on which you will implement all the discrimination models covered in class so far (nearest neighbors algorithm, naive Bayes, discriminant analysis)

## Part 1: Loading the dataset, exploratory analysis

```{r}
library(dplyr)
library(ggplot2)
library(tidyverse)
```

This dataset consists of measurements taken on several wines. For each wine sample, 11 characteristics were measured (acidity, density, pH, etc.) and after tasting, each received a score. For the purposes of this assignment, we will consider that each wine was assigned a class 1 (good quality wine) or class 0 (poor quality wine). The goal is to predict the class of a wine based on its 11 characteristics.

Questions:

1\. Load the dataset and analyze the summary of the table. What do you notice?

```{r}
Qualite_vin <- read.csv2("data/Donnees_qualite_vin.csv")

str(Qualite_vin)
```

On remet les colonnes dans le bon type

```{r}
Qualite_vin <- Qualite_vin %>%
  mutate(across(c(fixed.acidity, volatile.acidity, citric.acid, residual.sugar,
                  chlorides, free.sulfur.dioxide, total.sulfur.dioxide, density,
                  pH, sulphates, alcohol),as.numeric))
```

```{r}
Qualite_vin$quality <- as.character(Qualite_vin$quality)
```

On vérifie s'il y a des doublons

```{r}
sum(duplicated(Qualite_vin))
```

```{r}
Qualite_vin %>%
  filter(duplicated(.))
```

Il y a beaucoup de doublons, on les retire alors

```{r}
Qualite_vin <- Qualite_vin %>%
  distinct()
```

On passe maintenant à l'analyse

```{r}
summary(Qualite_vin)
```

2\. Using a preliminary visual exploration of the data, identify the characteristics that appear at first glance to most influence the quality of a wine.

```{r}
ggplot(Qualite_vin, aes(x = quality, y = alcohol, fill = quality)) +
  geom_boxplot() +
  labs(title = "Distribution de l'alcool selon la qualité du vin",
       x = "Qualité (0 = mauvais, 1 = bon)", y = "Alcool") +
  theme_minimal()

```

```{r}
vin_long <- Qualite_vin %>%
  pivot_longer(cols = -quality, names_to = "variable", values_to = "valeur")

ggplot(vin_long, aes(x = quality, y = valeur, fill = quality)) +
  geom_boxplot() +
  facet_wrap(~ variable, scales = "free_y") +   # une échelle libre par variable
  labs(title = "Distribution de chaque caractéristique selon la qualité du vin",
       x = "Qualité", y = "Valeur") +
  theme_minimal() +
  theme(legend.position = "none")

```
